Communication in Agile Inter-team Development

### 2019-02-04 Draft/Parts/Introduction.tex
Many minor changes. Mainly for length.

%AM: What is the problem?
%introduce the need to scale to multiple agile teams
Agile approaches are popular in software development for many reasons.
However, the structure of an agile process does not lend itself to large teams and is not suitable for large projects \cite{Rubin12}.
In complex software, situations arise where multiple agile teams need to work together to achieve a common goal.
%thesis statement
When scaling to multiple teams, the most important consideration is enabling effective communication, both personal and architectural, which is the focus of this paper.
%
For an example of a company developing a complex piece of software, consider Spotify AB who has over 30 teams \cite{kniberg12}.
Spotify is a streaming music client which supports media, commercial, and social features.
The basic element of development organization at Spotify is the Squad, analogous to a Scrum team.
%Each Squad is the autonomous developer, with a localized working area, of a particular sub-system in the product.
Squads in related areas are grouped together into Tribes. 
The overall structure is illustrated in Figure~\ref{fig:spotify_structure}.
\begin{figure}[h]
  \includegraphics[width=\linewidth]{images/kniberg12_structure.png}
  \caption{Agile team organization at Spotify \cite{kniberg12}.}
  \label{fig:spotify_structure}
\end{figure}

%AM: Why is it interesting/important?
%introduce agile
As outlined in the Agile Manifesto \cite{beck2001agile}, an agile process gives individuals and interactions value over processes and tools.
An agile team should contain all that is required for them to do their work, thus interactions are intensely intra-team.
However, diverse feature sets such as required in the Spotify application necessitate many skill sets and substantial manpower.
By using a multi-team model, large projects can benefit from agility, provided that the process is adapted.
%personal communication
Multiple teams increases the importance of communication, from simple interpersonal communication to system specification. 
The agile approach of interacting in-person can fail due to reasons such as the vertical structure of company organization and layers of middle managers \cite{dzone_article}.
The large number of intermediaries can cause an agile process to breakdown as it takes longer than usual for messages to reach their destination than can afforded.
%arch communication
In an agile process, where Big Design Up Front may typically be avoided, the move to multiple teams also motivates a need to communicate design decisions.
%todo: component boundaries?
For successful integration of technological components, system elements must have well defined functional specifications, and interfaces.

@@

%outline of rest of paper
This document is organized as follows: Introduction, Communication Concerns, Supporting Communication in Inter-team Development, and Conclusion. 
In Section ~\ref{sec:spt_ex}, we discuss example communication concerns.
In Section ~\ref{sec:prop_appro}, we present some process adaptations that can help resolve issues.
Lastly, Section ~\ref{sec:conclusion} concludes with an overview of important aspects of inter-team agile development.

>>>>

%AM: What is the problem?
%introduce the need to scale to multiple agile teams
Agile approaches are popular in software development for many reasons such as adaptability towards change, early customer satisfaction, and increase in team effectiveness with time.
However, the concept of an agile process does not lend itself to large teams and is not suitable for large projects \cite{Rubin12}.
In complex software, situations arise where multiple agile teams need to work together to achieve a common goal.
%thesis statement
When scaling to multiple teams, the most important consideration is enabling effective communication, both personal and architectural, which is the focus of this paper.
%
For an example of a company developing a complex piece of software, consider Spotify AB who has over 30 teams \cite{kniberg12}.
Spotify is a streaming music client which supports media, commercial, and social features.
The basic element of development organization at Spotify is the Squad, analogous to a Scrum team.
%Each Squad is the autonomous developer, with a localized working area, of a particular sub-system in the product.
Squads in related areas are grouped together into Tribes. 
The overall structure is illustrated in Figure~\ref{fig:spotify_structure}.
\begin{figure}[h]
  \includegraphics[width=\linewidth]{images/kniberg12_structure.png}
  \caption{Agile team organization at Spotify \cite{kniberg12}.}
  \label{fig:spotify_structure}
\end{figure}

%AM: Why is it interesting/important?
%introduce agile
As outlined in the Agile Manifesto \cite{beck2001agile}, an agile process gives individuals and interactions value over processes and tools.
An agile team should contain all that is required for them to do their work, thus interactions are intensely intra-team.
However, diverse feature sets such as required in the Spotify application necessitate many skill sets and substantial manpower.
By using a multi-team model, large projects can benefit from agility, provided that the process is adapted.
%personal communication
Multiple teams increases the importance of communication, from interpersonal to system specification. 
The agile approach of interacting in-person can fail due to reasons such as the vertical nature of a company and layers of middle managers \cite{dzone_article}.
The large number of intermediaries can cause an agile process to breakdown as it takes longer for messages to reach their destination than can be afforded.
%arch communication
In agile, where Big Design Up Front may be avoided, the move to multiple teams also motivates a need to communicate design. %decisions
%todo: component boundaries?
For successful integration of components, system elements must have well defined functional specifications and interfaces.

@@

%outline of rest of paper
This document is organized as follows: Introduction, Communication Concerns, Supporting Communication in Inter-team Development, and Conclusion. 
In Section ~\ref{sec:spt_ex}, we discuss communication issues.
In Section ~\ref{sec:prop_appro}, we present process adaptations that can help resolve issues.
Section ~\ref{sec:conclusion} concludes with an overview of key aspects of inter-team agile development.

### 2019-02-04 Draft/Parts/Main.tex
Many minor changes. Mainly for length.

\section{Communication Concerns}\label{sec:spt_ex} 
%overview
A fundamental issue in tasking teams with separate goals is that teams can become isolated from other teams pursuing similar or related goals.
% problem
Consider a scenario at Spotify \cite{kniberg12}: a tester in a particular Squad invested time to solve a particular problem, which was also faced by another tester on a different team. Without communication, these team members would perform redundant work.
To address this problem Spotify defines Chapters, which are crosscutting groups which gather individuals with similar roles across different Squads. 
At Spotify, offices are laid out so that the Squads making up a Tribe are spatially close \cite{kniberg12}.
This leads to an environment where an informal exchange of information on the Tribe's work will occur.
%Although larger than Squads (7-10 members), Tribes are limited in size (< 100) to ensure smooth communication.
%todo: any more issues? Dunbar

Consistency becomes an issue when expanding to multiple teams.
The product owner in an agile process model is the person with the project vision, and with a small team, they can clearly communicate it to every team member.
Members can even query the product owner directly when in doubt.
However, in large projects where multiple teams are working together, it is practically impossible for the owner to continue what they did in a small team.
This motivates a need for a formal definition of their vision as it gives teams something to consult to when encountering a design decision.
Without a functional specification, teams will be unable to produce components which can be easily integrated.

@@

%Changes in standups
One critical aspect of an agile process is daily standups which are a platform for letting each team member know the work being done in other parts of the team and correlate it with the work they are performing.
This uncovers differences in understanding early in development and prevents last minute discovery of discrepancies.
In the case of multiple agile teams this problem is compounded as a stand up is a closed activity of the team itself, thus preventing other teams from knowing the results or discussions of each other.
This can be resolved by having a representative of each team be present in a standup, e.g., a Scrum of Scrums, where typical stand up questions are given but reframed for a team context (e.g., "What will my team do before we meet again.." \cite{Rubin12}).

%Changes to Product Owner
Although the product owner is responsible for agile teams under his supervision, projects with multiple agile teams may have multiple product owners, and so over time the vision of the owners may drift and lead the development in different directions.
This can be limited by having regular meetings of product owners where the scope and vision of the project can be synchronized.
%This can be a bi-weekly or monthly meeting depending on the size of the project.

@@

\subsection{Technology Improvements} \label{sec:tech_impv}
Selecting an appropriate technology stack plays an important role in enabling communication and task management so that teams can work together efficiently.
%The following are areas where technology can enable a more efficient agile process.

%Communication Tools
Since agile focuses on interactions, like face-to-face conversation, video conferencing tools can lower the bar to inter-team interactions.
They also make communication real-time, reducing delay due to time spent communicating ideas across teams.
%Integration Tools
Continuous integration tools can go a long way in identifying inter-team problems early in development
as it reveals the impact of changes on the entire project.

@@

System design or architecture acts as a common vision for not only the teams but also for the communication between all the product owners and thus is the theme or topic for most of the inter-team communication that happens in a multi-team agile environment.
Being such a frequent piece of discussion, it needs to be as formal and as clearly defined as possible as it is the system design that acts as a “deadlock breaker in decisions” \cite{architecureRole_article} for situations where teams disagree on a particular design path, and ensures that components will be constructed so they can be integrated into the whole.
%non specification fix
A second approach addresses this issue from an organizational standpoint by defining a "system owner" role \cite{kniberg12}.
This role is more casual than an architect role, and focuses on defining a "go-to" person who can maintain a long term stewardship over a sub-system.

In addition to specification, proper architecture can also be used to enable team agility.
As discussed by Parnas \cite{Parnas72}, there are two general approaches to decomposing systems: 1) compartmentalizing a computational process and, 2) focusing on information hiding.
The latter approach is ideal for agile development since it defines system components in terms of hiding design decisions, which empowers individual teams to design their own solution.

Another way in which a formal architecture description helps in mitigating the communication problems is by minimizing the need for inter-team communication as a good design document is clear enough for a developer to incorporate the design in the system without the need of any other document.
Thus, a good design documents sometimes acts as a proxy for product owner when it comes to detailing the requirements of the customer.

>>>>

\section{Communication Concerns}\label{sec:spt_ex} 
%overview
A fundamental issue in tasking teams with separate goals is that teams can become isolated from other teams pursuing similar or related goals.
% problem
Consider a scenario at Spotify \cite{kniberg12}: a tester in a particular Squad invested time to solve a problem, which was also faced by another tester on a different team. Without communication, members would perform redundant work.
To address this problem Spotify defines Chapters, which are crosscutting groups which gather individuals with similar roles across different Squads. 
At Spotify, offices are laid out so that the Squads making up a Tribe are spatially close \cite{kniberg12}.
This leads to an environment where an informal exchange of information on the Tribe's work will occur.
%Although larger than Squads (7-10 members), Tribes are limited in size (< 100) to ensure smooth communication.
%todo: any more issues? Dunbar

Consistency becomes an issue when expanding to multiple teams.
The product owner in an agile process model is the person with the project vision.
In a small team, they can clearly communicate it to every team member, and
members can even query the product owner.
However, in large projects with multiple teams, it is practically impossible for the owner to continue what they did in a small team.
This motivates a need for a formal definition of their vision as it gives teams something to consult to when encountering a design decision.
Without a functional specification, teams will be unable to produce components which can be easily integrated.

@@

%Changes in standups
One aspect of an agile process is daily standups which are a platform for letting each team member know the work being done by other team members. % and correlate it with their work.
This uncovers differences in understanding early and prevents last minute discovery of discrepancies.
For multiple teams, this problem is compounded as a stand up is a closed activity, preventing other teams from knowing the discussions of each other.
This can be resolved by having a representative from each team be present in a standup, e.g., a Scrum of Scrums, where stand up questions are given but reframed for a team \cite{Rubin12}.

%Changes to Product Owner
Although the product owner is responsible for agile teams under his supervision, projects with multiple agile teams may have multiple product owners, and over time the vision of the owners may drift and lead in different directions.
This can be limited by regular meetings of product owners where the scope and vision of the project can be synchronized.

@@

\subsection{Technology Improvements} \label{sec:tech_impv}
Selecting appropriate technology plays an important role in enabling communication and task management so that teams can work together efficiently.
%
%Communication Tools
Since agile focuses on interactions, like face-to-face conversation, conferencing tools can lower the bar to inter-team interactions.
They make communication real-time, reducing delay in communicating ideas across teams.
%Integration Tools
Continuous integration tools can go a long way in identifying inter-team problems early in development as it reveals the impact of changes on the entire project.

@@

System architecture acts as a common vision for not only the teams but also for communication between the product owners, and is the main topic for most of the inter-team communication that occurs in a multi-team environment.
%Being such a frequent piece of discussion
It needs to be as formal, and as clearly defined as possible, as it is the system design that acts as a “deadlock breaker in decisions” \cite{architecureRole_article} for situations where teams disagree on a particular design path, and ensures that components will be constructed so they can be integrated into the whole.
%
A formal architecture description also helps to mitigate communication problems by reducing the need for inter-team communication as a good design document provides enough guidance for a developer to incorporate their work into the system.
Thus, a good design document sometimes acts as a proxy for a product owner when it comes to detailing the requirements of the customer.
%non specification fix
A second approach addresses this issue from an organizational standpoint by defining a "system owner" role \cite{kniberg12}.
This role is more casual than an architect role, and focuses on defining a "go-to" person who can maintain a long term stewardship over a sub-system.

Architecture can also be used to enable team agility.
As discussed by Parnas \cite{Parnas72}, there are two general approaches to decomposing systems: 1) compartmentalizing a computational process and, 2) focusing on information hiding.
The latter approach is ideal for agile development since it defines components in terms of hiding design decisions, which empowers teams to design their own solution.
